---
title: "Targets_implementation"
author: "Rostyslav Vyuha"
date: "April 14, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Main Export Functions

### Verify Targets

The purpose of this function would be to verify the correct order names and arguments in the passed targets.
This is done by comparing it against the read in modules

  The function would contain 2 arguments **targets_source** and **modules_path**.
The **targets_source** would either be the tar_target list or in most cases the path to the _targets.R
The **modules_path** would be the file path to the modules_map.csv

This function would not make any changes to the _targets.R file or the list itself
and would simply output warnings and a boolean representing the validity of the targets with the module.

#### List of warnings
- error when a module step is missing
- error when module steps are out of order
- error when module step contains wrong arguments

#### Example function usage

```{r, echo=FALSE}
verify_targets(targets_source = "/assets/specs/targets/depression_targets.R", modules_path = "/assets/specs/targets/modules_map.csv")
```
This returns TRUE if depression_targets.R contains everything inside modules_map and in correct order with correct arguments

### Run bllflow Targets

This function would be responsible for running the targets with arguments filled in by the bllflow object.

Excluding the arguments mentioned above this function would contain 2 arguments **targets_source** and **bllflow_object**.
The **targets_source** would be identical to the one to verify targets.
The **bllflow_object** would be the bllflow object created upon config initialization with mandatory checks for modules.csv and variables.csv as well as a present working_data.

The function would first run verify targets to confirm correct order and presence of steps. Then it would modify the tar_targets arguments to reflect their true value rather then the shorthand (roles).
Once the tar_targets were modified accordingly the _targets.R file is written and tar_make() is executed, letting targets handle the returns and the pipeline

#### Example function usage

```{r, echo=FALSE}
run_bllflow_targets(targets_source = "/assets/specs/targets/depression_targets.R", bllflow_object = hui_object)
```

This would create a _targets.R in base package directory using the targets found at targets_source.
It would essentially be a copy and paste except for the Special Arguments which would be populated using the bllflow_object

### Create _targets tepmlate

This function would be responsible for creating the basic bllflow_targets.R file which would only be populated by the steps in modules.csv

The function would once again contain only 2 arguments **target_path** and **modules_path**

The function would utilize the shorthand(roles) notation when writing the functions for ease of use for the analyst

#### Example function usage

```{r, echo=FALSE}
create_targets_tepmlate(target_path = "/assets/specs/targets/depression_targets.R", modules_path = "/assets/specs/targets/modules_map.csv")
```

This would create a barebones depression_targets.R with only things found in the passed modules

### Create _targets list

This functions would be responsible for creating a list containing tar_target objects.

The function would accept 1 mandatory arguments **modules_path** and one optional argument **target_path**.
If a **target_path** is supplied the existing tar_targets list is read in and appended and verified before being returned, if no **target_path** is supplied a barebone template type list is created from the modules_map

#### Example function usage

```{r, echo=FALSE}
create_targets_list(modules_path = "/assets/specs/targets/modules_map.csv")
```

This would create a barebones tar_targets list with only things found in the passed modules

# Contents of modules.csv

### Step_id

The step_id column must contain a unique identifier for the step being performed.

### Step_function

The step_function column contains the name of the function being performed in this step. This must match a function name present in the environment during execution.

### Step_argument_name

The step_argument_name as the name implies contains the name of a single argument

### step_argument_value

The step_argument_value contains the value for a single argument that matches the name in step_argument_name

#### Special Arguments

*role* This would search for variables matching the role in variables.csv and be replaced with vecor of var names during run time.
*data* This would pass the object attached to the bllflow object inside the data list ie: bllflow$data[[<whats inside data>]], alternatively it can be a reference to data generated by a previous step.
*formula* This would create a left side = right side formula ie: formula[role["outcome"], role["predictor"], sep = "+"] would result in "outcome1 + outcome2 + outcome3 ~ predictor1 + predictor2 + predictor3"

### Step_description

This column contains the step description this is used to populate comments in template creation function. It should contain a helpful description of what this step is responsible for.

### Step_order

This column contains the order in which steps should be executed

# Contents of modules_map

### Module_Name

The name of the module being included

### Module_Path

The relative path to the module being loaded

### Module_description

The module being ran

### Module_order

The order in which modules are ran

# Example function usage

## Verify_targets


## run_bllflow_targets






