---
title: "Reference files - variable labels and metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reference files - variable labels and metadataa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
Data 
and models are meaningless without good metadata.  Metadata should be 
independently understandable; machine actionable; discoverable, dynamic, 
interactive. 
**bllFlow** uses consistent labels and metadata throughout the workflow. We explain why in our accompanying [document]. tl;dr [Nature Videos](https://youtu.be/j7K3s_vi_1Y)

## `variable-metadata`
`variable-metadata` is the data file that describes labels and metadata in bblFlow. Add to your own `cat-met-labels` or make a [pull request](https://github.com/big-Life-Lab/bllflow/pulls). 

Labels and metadata are used in five key areas of research reporting:

1) **Data cleaning and variable transformation** -- such as applying truncation rules, centering, etc.
1) **Aggregated results** -- such as _Table 1 - Characteristics of study population_.
1) **Model description** -- model coefficeints.
1) **Summary statistics** -- metrics model performance.
1) **Validation data** -- example dataset to verify algorithm scoring (for predictive models).

`cat-met-labels`

```{r, echo=FALSE, results='asis'}
library(DT)

dt <- read.csv(system.file("extdata", "summary_variables.csv", package = "bllFlow"))
datatable(dt, options = list(pageLength = 5))
```
