---
title: "1 - Specifying your model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1 - Specifying your model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction
The Model Specification Workbook (MSW) is used to specify your model. The MSW is 
a series of four worksheets (CSV files) that describe different model components. 
You can use **bllflow** without a MSW but we recommend using the 
`variables` and `variable_details` worksheets. 
 
## The four MSW worksheets

The Model Specification Workbook (MSW) is the central document creating your 
model using _bllflow_. 
Worksheets within the MSW contain both the names of the starting variables in the original study 
data and the final variables, if variable are transformed. 
- **variables.csv** - Each row in the `variables.csv` file is a variable in your model. 
The columns include variable labels, type, and other information such as which 
variables are transformed (centered, have dummy variables, restrictive cubic 
splines, have interaction terms, etc.).
- **variablesCetails.csv** - includes the details of each variable, including all 
category (factor) levels and labels. Also included are  specific variable transformation 
rules.
- **modelCescription.csv** - includes information such as the model name, outcome, 
model, creator(s), date and data set names.
- **tableCescription.csv** - includes specifications for how to create Table 1 
and other tables.

There are two approaches to creating the MSW. We usually start the MSW as a CSV 
file to facilate collaboration -- and this approached is used throughout the 
vignettes. Alternatively, you can create the MSW as an R dataframe. 
## Getting stated with the Model Specification Worksheets.

The first step in _bllflow_ is to identify all starting and final variables for 
your model in the MSW. We emphasize pre-specified analyses but 
variables can be added as you perform study. We highly recommend any additional 
variables and transformation be added to the MSW to ensure reproducibilty and 
transparency. The MSW is a record of how you created your model and what analyses
 you performed. As well, _bllflow_ uses metadata throughout the workflow. 
You have your study data. Great. The first step is specifying which variables you 
need in your model as well as variables for study cohort creation, data cleaning and variable transformation. 

The second step is to add variable labels and other metadata to the MSW that 
provides instructions for data cleaing and transformation. 
_bllflow_ has example Model Specification Work sheets for the `pbc` data and the 
vignette shows how to recreate a survival model for [primary biliary cirrhosis](https://aasldpubs.onlinelibrary.wiley.com/doi/pdf/10.1002/hep.1840100102). However, the specific work for many model types. 

*bllFlow* supports importing metadata from:
## Examples of the Model Specification Worksheet
```{r} 
library(DT)

- DDI (xml) documents;
- variable lablels in study dataframe as `attr` label (using hmisc, sjlabelled or
similar packages); or,
- manually added added to the MSW.
dt <- read.csv(system.file("extdata", "PBC/PBC-variables.csv", package = "bllFlow"))
datatable(dt, options = list(pageLength = 5))
```

## Worksheets 
**bblFlow** has four worksheets within the MSW:
## Importing the MSW worksheets
```{r}
library(survival)
data(pbc) # pbc survival data

1) `modelDescription` --- the name of the model, date created and other 
information about the study.
1) `variables` --- all model variables, including data cleaning and 
transformations. 
1) `variableDetails` --- information on factors (categories) and how to transform
 final variables from their starting variables. 
1) `summaryVariables` -- variables that are used in model reporting, such as 
for `Table 1`. 
# Read variables and variableDetails sheet for the PBC model
variablesSheet <- read.csv(file.path(getwd(), 'bllFlow/extdata/PBC/PBC-variables.csv'))
variableDetailsSheet <- read.csv(file.path(getwd(), 'bllFlow/extdata/PBC/PBC-variable-details.csv'))

```
# pbc data for examples and vignettes
library(survival)
data(pbc)
```
```{r}
# Create a bllFlow R object for the PBC model using the above variables as args
library(bllFlow)
#' pbcModel <- BLLFlow(pbc, variablesSheet, variableDetailsSheet)

Read in the MSW and variable_details sheet for the PBC model
modelDescription <- read.csv(file.path(getwd(),..
variables <- read.csv(file.path(getwd(), 'bllFlow/extdata/PBC/PBC - variables.csv'))
variableDetails <- read.csv(file.path(getwd(), 'bllFlow/extdata/PBC/PBC - variable_details.csv'))
```

# Create a bllFlow R object for the PBC model using the above variables as args
library(bllFlow)
pbcModel <- BLLFlow(pbc, variablesSheet, variableDetailsSheet)
```
**bllFlow** also supports the use of DDI documents. If available, bllFlow labels 
and other information to the `variables.csv` and `variablesCetails.csv`, as well
as perserving data information to support data provenance. 




