---
title: "Dummy coding"
author: "Carol Bennett"
date: "2020-10-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Unlike dichotomous or continuous variables, categorical variables (with more than two levels) cannot be entered into a regression equation as they are—instead, they need to be recoded into a series of variables which can be entered into the model. Dummy coding compares each level of a variable to the omitted (reference) level. It is a way to make the categorical variable into a series of dichotomous variables. For all but one of the levels of the categorical variable, a new variable will be created that has a value of one for each observation at that level and zero for all others.

### Example

Consider an education variable with four levels: less than high school, high school graduate, some post-secondary education, and post-secondary school graduate.

If we identify 'post-secondary school graduate' as the reference group, new variables (e.g., 'less_than_hs', 'hs_grad', 'some_ps') will be created for the other levels. 

```{r read variables.csv, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(DT)
library(knitr)

x <- data.frame("Level" = c("less than high school", "high school graduate", "some post-secconday education", "post-secondary school graduate"), "less_than_hs" = c(1,0,0,0), "hs_grad" = c(0,1,0,0), "some_ps" = c(0,0,1,0), stringsAsFactors = TRUE)

knitr::kable(x)
```

The new variable 'less_than_hs' will have a value of 1 for observations with 'less than high school' education level and 0 for all other observations. The new variable 'hs_grad' will have a value of 1 for observations with 'high school graduate' education level and 0 for all other observations, and so on.

## Creating dummy variables with bllflow

