---
title: "7 - (predictive) Model deployment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{7 - (predictive) Model deployment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

This section is in progress. _bllflow_ use the Model Specification Workbook to provide instructions to generate:

- the PMML file that described how to calculate predictive algorithms.
- validation (testing) data.
- calibratin data.
```{r}
library(survival) # for pbc data
data(pbc)
library(bllflow)
pbcDDI <- ReadDDI(file.path(getwd(), "../inst/extdata"), "pbcDDI.xml")
variables <- read.csv(file.path(getwd(), '../inst/extdata/PBC-variables.csv'))
variableDetails <- read.csv(file.path(getwd(), '../inst/extdata/PBC-variableDetails.csv'))

# create a BLLFlow object and add labels.
pbcModel <- BLLFlow(pbc, variables, variableDetails, pbcDDI)
WriteDDIPopulatedMSW(pbcModel, "../inst/extdata/", "newMSWvariableDetails.csv")
WriteDDIPopulatedMSW("../inst/extdata/", 
                  "newMSWvariableDetails.csv", 
                  "nonBLLPopulated.csv")
 
```
See the following examples of how `bllflow` is used to report and deploy models.

  * Algorithms for indivdual use at [Project Big Life](https://www.projectbiglife.ca)
  * Application Program Interface [(API)](https://ottawa-mhealth.github.io/pbl-calculator-engine-docs/)
  * Algorithm deployment for population planning - Coming September 2019.