---
title: "Helper and utility functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Healper and utility functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{survival}
  % \VignetteDepends{dplyr}
  % \VignetteDepends{magrittr}
  
---

## Introduction

## Metadata

_bllflow_ supports metadata from DDI and from the R packages hmisc and rjlabelled.

### DDI metadata
The [Data Documentation Intitiatve](https://www.ddialliance.org) (DDI) is is an 
international standard for describing the data produced by surveys and other 
observational methods in the social, behavioral, economic, and health sciences. 
There is metadata for variables labels, but also important data provenance information.

Get the description of the study data from DDI and add the description to 
_bllflow_ object

```{r Load libraries, warning=TRUE, include=FALSE}
library(survival) # for pbc data
data(pbc)
library(bllflow)
```

### Example 1: Read the DDI xml file for your data.

```{r}
pbcDDI <- ReadDDI(file.path(getwd(), "../inst/extdata"), "pbcDDI.xml")
```
Your metadata is now stored in the `pbcDDI` object in two formats:

1) Variable and format labels can be accessed in `pbcDDI$variableMetaData`. 
All DDI metadata is brought into the `pbcDDI` object. For example, variable labels can 
be accessed as follows:
```{r}
str(pbcDDI$variableMetaData$varlab) # variable labels
str(pbcDDI$variableMetaData$vallab) # value labels for categorial variables
```
`variableMetaData` is generated using the `DDIwR` package.

2) All metadata from the DDI document `pbcDDI$ddiObject`. Careful
Get the varialbe metadata for variableDetails sheet from DDI and add 
that the bllFlow MSW-DDI object. Careful printing: there can be a lot
of information! And there can be a lot of nested lists.

Get the name of the data.
```{r}
cat("Dataset name: \n")
pbcDDI$ddiObject$codeBook$docDscr$citation$titlStmt$titl
```

#### Example 2: Use DDI to add variable labels and other information for your model variables.

The MSW `variables` and `variableDetail.csv` file contains the variables in your model. 
Use the `BLLFlow()` function with the MSW files as attributes to add labels to the 
BLLFlow object, (`variables` and `variableDetails` and  the DDI object (`pbcDDI`).
```{r}
# read the MSW files
variables <- read.csv(file.path(getwd(), '../inst/extdata/PBC-variables.csv'))
variableDetails <- read.csv(file.path(getwd(), '../inst/extdata/PBC-variableDetails.csv'))

# create a BLLFlow object and add labels.
pbcModel <- BLLFlow(pbc, variables, variableDetails, pbcDDI)
```
Metadata is added to `variables` and `variableDetails`. If labels and other metadata 
already were in the files, the DDI metadata is added to `startLabel`, `startType`, 
`catStartValue`, `catStartLabel`, `startLow`, `startHigh` --- if that data is in 
the DDI file. 

```{r}
cat("Variable labels in the original MSW\n")
variables$label

cat("\nNo variable labels from DDI in variable details\n")
variableDetails$label

cat("\nDDI variable lables added to variableDetailsWithDDI\n")
pbcModel$populatedVariableDetails$label
```

#### Example 3: Write MSW_DDI to MSW.csv
Use `WriteDDIPopulatedMSW()` to create a MSW `variable details` CSV file. 
(change to new name `variableDetailsWithDDI()`) using `variableDetailsWithDDI()`. 
Then export using `WriteDDIPopulatedMSW()` as a CSV file to help further develop 
your study protocol. 

`WriteDDIPopulatedMSW()` is a handy function at the beginning of your study. 
First, identify thes ariable you need in your study. Then create a 'bane bones' 
MSW `variables` CSV sheet, import into R and `BLLFlow()` to add additional 
information from the DDI reference file such as labels, variable types, categories. 

Alternatively, 
1) directly add metadata to variables in a CSV file (see general utility 
functions in *Example X*); or, 
2) export a CSV with metadata from an R variable list and a DDI file.

```{r}
WriteDDIPopulatedMSW(pbcModel, "../inst/extdata/", "newMSWvariableDetails.csv")
```
Creates a directory and file (if they don't already exists). `newMSWvariableDetails.csv` 
will be overwritten if the file already exists. `newMSWvariableDetails.csv` is 
created from `variableDetailsWithDDI`. First create `variableDetailsWithDDI` 
with either `BLLFLow()` or `GetDDIVariables()`.

Create a new file (`newName.csv`) directly from an existing variableDetails.csv.
```{r}
WriteDDIPopulatedMSW(pbcDDI, "../inst/extdata/", 
                  "newMSWvariableDetails.csv", 
                  "nonBLLPopulated.csv")
```

#### Example 4: Update BBLFLow with new variables

Do we need or want this? Rusty, can you write in a vignette?
Replaces the models msw with all or either variables or variableDetails. 
Passing variableDetails also updates the populatedVariableDetails

```{r}
pbcModel <- UpdateMSW(pbcModel, variables, variableDetails)
```

#### Example 5. More general DDI utily functions 
Described previously,`ReadDDI()` returns a dataframe with all DDI metadata from 
a DDI.xml file. 

There two additional DDI utility functions that return the two main parts of the 
DDI file.

`GetDDIDescription()` returns a dataframe with DDI 'header' information from a DDI file. 
The header information includes the dataset ID, name, creatation date and other 
important information to preserve the provenance of your study data.

`GetDDIVariables()` returns a dataframe with variable DDI metadata for a list of 
variables. DDI variable metadata is typical codebook information such as labels, 
types, valid and invalid categories, category labels, and descriptive statistics, 
such as number valid responses, missing responses, minimum, maximum and mean values. 

The _bllflow_ workflow creates an ongoing, updated codebook as you perform your 
study. The oringial data DDI metadata is the starting point for keeping the updated 
codebook. The _bllflow_ then creates and modifies an new instance of the codebook 
as you perform analyses. In addition, a log is created that describes the data 
cleaning and data transformation steps. 

getDDIDescription(DDIPath = ddiPath, DDIFile = ddiFile)
# same as calling BLLFlow(pbc, getDDI = c(ddiPath, ddiFile, "description"), but 
would not add the info to the bllFlow object.

```{r}

pbcDDI <- ReadDDI(file.path(getwd(), "../inst/extdata"), 
                "pbcDDI.xml")
# all the DDI metadata

varsForPbc <- GetDDIVariables(pbcDDI, c("age", "sex"))
# all the variable metadata (but no study description infomation)

descrForPbc <- GetDDIDescription(pbcDDI)
# all the study description metadata (but no variable infomation)
```