---
title: "Helper and utility functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Healper and utility functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{survival}
  % \VignetteDepends{dplyr}
  % \VignetteDepends{magrittr}
  
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
**bllFlow** supports pre-specifying model development as described by Harrell 
and others.[ref]


## Make an event date
Our teams often work with health data in secure environments where privacy is 
paramount. This means we cannot share data. ...
```{r make event date}
library(bllFlow)
library(survival)
library(dplyr)
library(magrittr)
library(Hmisc)

df <- system.file("extdata", "RESPECT-EOL_validation.csv", package = "bllFlow") %>%
  read.csv %>%
  as.data.frame

df_length <- nrow(df)

# make events
# df$tt_event <- make_event_dates(df_length, df$risk, 1825, label = "FollowUp", units = 'days')
# label(df$tt_event) #check label
# 
# # check to see if events have been added
# table(df$tt_event)
# 
# # check to see if df$tt_event is about the same as df$risk
# mean(df$risk)
# nrow(df[df$tt_event < 1825,]) / nrow(df)
# 
#  # create event variable
# df$event <- ifelse(df$tt_event < 1825, 1, 0)
# 
# # Pick an available survial time less than 1825 for evaluation
# # (This is only neccessary due to current limitations of the test dataset)
# a <- sort(df$tt_event) %>% unique
# toe <- tail(a, 2)[-2]
```

## Filter table 1
Assessing model performance is an important aspect to health care research. In order to simplify this post modeling step a helper function was created to filter table one data. This function is intrinsic for visual inspection model assessments (ie. Calibration plots). Filter_table_1 can filter data based on conditions given from either, all, or any combination of the three groupby variables typically seen in our algorithm development (group_by_sex, group_by_1, and group_by_2) 
```{r filter table one, echo=TRUE}
library(bllFlow)

#load dataset
df <- system.file("extdata", "MPoRT-table-one.csv", package = "bllFlow") %>%
  read.csv %>%
  as.data.frame

#Check variables
head(df, num=1L)

#list of variables to keep and filter in subset of the dataset
var<-c("observed_risk_1_year", "predicted_risk_1_year", "group_by_sex", "group_by_1", "group_by_2","group_by_value_label_1") 

#Examples
filter_table_1(df, var, groupby_sex="Male", groupby_1 ="smoking", groupby_2="packyears")

filter_table_1(df, var, groupby_sex="Female", groupby_1="age")

filter_table_1(df, var, groupby_1="age")


```

