---
title: "Helper and utility functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Healper and utility functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{survival}
  % \VignetteDepends{dplyr}
  % \VignetteDepends{magrittr}
  
---

## Introduction

### Metadata

_bllFlow_ supports metadata from DDI and from the R packages hmisc and rjlabelled.

#### DDI metadata
The [Data Documentation Intitiatve](https://www.ddialliance.org) (DDI) is is an 
international standard for describing the data produced by surveys and other 
observational methods in the social, behavioral, economic, and health sciences.

```
Get the description of the study data from DDI and add the description to 
bllFlow object

```{r}
library(survival)
data(pbc)
library(bllFlow)
```

Example 1: First, read the DDI xml file for your data.
```{r}
pbcDDI <- ReadDDI(file.path(getwd(), "inst/extdata/test-folder/DDI"), "pbcDDI.xml")
```

Rusty -  it shows variable and value label lists
All DDI metadata is brought into the `pbcDDI` object. For example, variable labels can be accessed as follows:
```{r}
pbcDDI$variableMetaData
```

Get the varialbe metadata for variableDetails sheet from DDI and add 
that the bllFlow MSW-DDI object.
```{r}
variables <- read.csv(file.path(getwd(), 'inst/extdata/PBC/PBC - variables.csv'))
variableDetails <- read.csv(file.path(getwd(), 'inst/extdata/test-folder/DDI/PBC - variableDetails.csv'))

pbcModel <- BLLFlow(pbc, variables, variableDetails, pbcDDI)
```

Example 2: Write MSW-DDI to MSW.csv
```{r}
WriteDDIPopulatedMSW(pbcModel, "inst/extdata/PBC/vignette-testing/", "testing.csv")
WriteDDIPopulatedMSW(pbcDDI, "inst/extdata/PBC/vignette-testing/", "testing.csv", "newName.csv")

```

# General utily functions 
(These functions could already be in DDIwR?). Four functions.

getDDIdescription(DDIPath = ddiPath, DDIFile = ddiFile)
# same as calling BLLFlow(pbc, getDDI = c(ddiPath, ddiFile, "description"), but 
would not add the info to the bllFlow object.

Example 3: general utility.
```{r}
pbcDDI <- ReadDDI(file.path(getwd(), "inst/extdata/test-folder/DDI"), "pbcDDI.xml")
CCHS_DDI <- ReadDDI(file.path(getwd(), "inst/extdata/test-folder/DDI"), "cchs-82M0013-E-2013-2014-Annual-component.xml")

varsForCCHS <- GetDDIVariables(CCHS_DDI, c("DHHGAGE", "DHH_SEX"))
```
This needs a newPoRT-MSW file to be there in order to populate from ddi otherwise it will not work
```{r}
WriteDDIPopulatedMSW(CCHS_DDI, "inst/extdata/PBC/vignette-testing/", "newPoRT-MSW", "newName.csv")
```
creates a new MSW-variableDetails sheet called "newPort-variableDetails.csv" 
with one variable: DHHAGE

Example 4: Should we directly import the DDI.xml file within the BLLFlow() ojbect, as currently written, or first have people invoke the general utility? The latter approach would 
be more consistent with how we calling `variables` and `variableDetails`
```{r}
pbcDDI <- ReadDDI(file.path(getwd(), "inst/extdata/test-folder/DDI"), "pbcDDI.xml")
pbcModel <- BLLFlow(pbc, variables, variableDetails, pbcDDI)
```
