---
title: "pbc-recode-with-table"
author: "Rostyslav Vyuha"
date: "October 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import pbc msw

```{r, results='hide',message=FALSE}
library(readr)
varDetails <- read_csv("../inst/extdata/PBC-variableDetails.csv")
variables <- read_csv("../inst/extdata/PBC-variables.csv")
```

## Import the pbc data

```{r, results='hide',message=FALSE}
library(survival)
data("pbc")
```

## Recode data with the msw

```{r, results='hide',message=FALSE}
library(bllflow)
recodedPBC <- RecWTable(dataSource = pbc, variableDetails = varDetails, datasetName = "survival-pbc")
```
## Compare the recode

```{r}
print(pbc)
print(recodedPBC[,c("TIME","STATUS","TRT","AGE","SEX","ASCITES","HEPATO","SPIDERS","EDEMA","BILI","CHOL","ALBUMIN","COPPER","ALK_PHOS","AST","TRIG","PLATELET","PROTIME","STAGE")])
```
## Write custom function for calculating age group

```{r}
CustomAgeGroupCalculation <- function(age){
  # Feel free to write ur own code this is an example
  return(trunc(age/10))
}
```

## Create a new variable that uses age group
I will use R code to create  one row for the new variable, then rowbind it with existing variable details feel free to change any elements u wish just make sure its consistent with the function arguments.

```{r}
newRow <- data.frame(variable = "Age_Group_Using_Custom_Variable", toType = "cont", databaseStart = "survival-pbc", variableStart = "DerivedVar::[AGE_GROUP, SPIDERS]", fromType = "cont", recTo = "Func::CustomRandomFunction", catLabel = "NA", catLabelLong ="NA", units = "NA", recFrom = "else", catStartLabel = NA, notes = NA, interval = NA)
varDetails <- rbind(varDetails,newRow)
```

## Create the other custom function

```{r}
CustomRandomFunction <- function(ageGroup, spiders) {
  # make sure to account for na since in R anything == NA equaites to NA and causes errors
  if (is.na(spiders)) {
  ageGroup <- ageGroup
  }
  else if (spiders == "1") {
  ageGroup = ageGroup + 5
  } else if (spiders == "2") {
  ageGroup = ageGroup - 5
  }
  
  return(ageGroup)
}
```

## Rerun the recode function to create new variable 
```{r}
recodedPBC <- RecWTable(dataSource = pbc, variableDetails = varDetails, datasetName = "survival-pbc")
print(recodedPBC)
```