---
title: "pbc-operations-test"
author: "Rostyslav Vyuha"
date: "July 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step one initial variable role specification
Iniside variable details specify which variables are predictors, used in creation of other vars, and which variables are the outcome(this is used to build the initial recipy) @Doug(perhaps we should move created variables into variables as well and move role specification to variables too?)

##Specify the modules
Inside the modules csv specify the functions you wish to use for each step

##Specify module effects
Inside the variables csv specify which variables are affected by which modules as well as the arguments for those modules for each variable

##Run the modules
Run the modules sequentially then procced to run checks on each modules as the analyst sees fit

```{r}
library(survival)
library(bllflow)
data(pbc)

#Import the ddi,variables,variableDetails,modules
variables <- read.csv(file.path(getwd(), '../inst/extdata/Meeting Version - Variables.csv'))
variableDetails <- read.csv(file.path(getwd(), '../inst/extdata/Meeting Version - variableDetails.csv'))
modules <- read.csv(file.path(getwd(), '../inst/extdata/Meeting Version - Modules.csv'))
ddi = pbcDDI <- ReadDDI(file.path(getwd(), "../inst/extdata"), 
                "pbcDDI.xml")

bllFlowObject <- BLLFlow()

# This updates the working data inside bllFlow to the data returned by RunModule as well as its current module sequence number
bllFlowObject <- RunModule(bllObject = bllFlowObject, moduleSequenceNumber = 1)
#The non bllFlow version
#Returns a list containing the working data, recipy log, as well as the module number as its elements
workingData <- RunModule(variables = variables, variableDetails = variableDetails, modules = modules, data = pbc, moduleSequenceNumber = 1)

# The analyst can then run required analysis on workingData$data to verify the module performed the steps they wished if not they can update module csv module 1 to run what they prefer to do in this step and rerun until they are satisfyid with its output

#CHECKS AND ANALYSIS
#rerun
workingData <- RunModule(variables = variables, variableDetails = variableDetails, modules = modules, data = pbc, moduleSequenceNumber = 1)

#Once the step is satisfied next module is ran with the first module output
workingData <- RunModule(variables = variables, variableDetails = variableDetails, modules = modules, data = workingData, moduleSequenceNumber = 2)

#This is then continued for as many default modules as there are created or however many the analyst wishes to run
#The working data/bllflow object will store a log for each module ran however will overwrite if the same module number is ran twice
```


