---
title: "Data Visualization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Healper and utility functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{plotly}
 
---

## Data Visualization


Visualizing data is an important aspect for any statistical modeling and research. For our purposes we have created 
two functions, one creates a basic calibration plot (predited versus actual), and the other an observed versus 
predicted bar chart scatter plot hybrid.



```{r echo=TRUE}
#load packages
library(bllFlow)
library(Rcurl)
library(plotly)

#load dataset
df <- read.csv("https://raw.githubusercontent.com/Big-Life-Lab/bllFlow/data-visualization/inst/extdata/MPoRT-table-one.csv?token=ApmTWVvRI4hGeBRlPLR6R3P6J73Po5Paks5cWzfswA%3D%3D", 
fileEncoding="UTF-8-BOM")

```

###Using helper functions in visualization process

In order to help simplify the visualization process a helper function was created to
filter the table one data frame.

*filter_table_1(data, vars, groupby_sex=NA, groupby_1=NA, groupby_2=NA)*


```{r filter_table_1, echo=TRUE}

col<-c("observed_risk_1_year", "predicted_risk_1_year", "group_by_sex", "group_by_1", "group_by_2","group_by_value_label_1") #columns to keep in subset of the dataset

# Example 1:
ex1<-filter_table_1(df, col, groupby_sex="Female", groupby_1="age")  #creates a subset of data with only the observations that meet the specified criteria  

#Example 2:
ex2<-filter_table_1(df, col, groupby_sex="Male", groupby_1="Drinks per week")

```


###Calibration Plot

Calibration plots are one of the richest and most utilized forms of data visualization. 
It is mostly used to visualize and evaluate goodness of fit (R Square).  The closure the
data points are to the 45â° diagonal line the stronger the goodness of fit of the model.
In terms of R square, if your R square was high all the points on the graph would be
close to the line. The lower the R square is the further away from the diagonal line
the points would lie, and the weaker the Goodness of fit for the model is.
Extremely distance points from the line represent possible outliers.

*calibration_plot(data, observed, predicted, error=0, title=NA)*

```{r Calibration Plot, echo=TRUE}
# Example One:
calibration_plot(ex1, "observed_risk_1_year", "predicted_risk_1_year", title="Canadian Female age")

# Example Two:
calibration_plot(ex2, "observed_risk_1_year", "predicted_risk_1_year", title="Drinks per Week for Canadian Males")
```

###Observed versus Predicted

Another way to visualize goodness of fit is with our observed vs predicted plot. 
Different from the more traditional observed vs predicted scatter plot, this graph
visualizes the observes in a bar chart and the predicted as overlaying points on the
bars. Like the calibration plots, the further away the predicted points are from the
observed bars the weaker the goodness of fit of the model would be. Extremely 
far off predicted points represent possible outliers.

*op_plot<-function(data, observed, predicted, title = NA, xlab = NA)*

```{r Observed vs Predicted, echo=TRUE}
# Example One:
op_plot(ex1, "observed_risk_1_year", "predicted_risk_1_year", title="Canadian Female age", xlab="Age")

# Example Two:
op_plot(ex2, "observed_risk_1_year", "predicted_risk_1_year", title="Drinks per Week for Canadian Males", xlab="Drinks per Week")

```

